Chasing \rainbows{} with the Habitable Worlds Observatory

### 2023-03-08 src/tex/ms.tex
[showyourwork] overleaf sync

\begin{figure}
   \centering
   \includegraphics[width=0.99\columnwidth]{figures/ellipse.pdf}
   \caption{
  Ellipse drawingIf you agree that this makes sense, I could try to add this to the yarn ball script by Matt? \todo{sophia: I have posted code in the slack to calculate beta min-max for randomly orientated eccentric orbits. I don't know how to add it to the code we already have.}}
    \label{fig:ellipse}
%    \script{plot-orb-grid.py}
\end{figure}

@@

\subsection{Simulating contrast curves}
\textbf{Kim and David}
%
A second parameter that is crucial for detecting features like rainbows and ocean glint on Earth-like planets is the contrast, i.e. the relative flux of the reflected light from the planet divided by the stellar flux. 
%
While this is not the main focus of this paper, it is important to consider that the contrast and polarised contrast are not constant as function of scattering phase angles.
%
This is due to a combination of the changing illumination fraction of the exoplanets Earth facing surface and scattering effects.
%
Here, we aim to present some simplified examples of the impact of the scattering angle dependent contrast on the detectability of rainbow and glint features. 
%
We note that these calculations are by no means complete or exact, given the many variables that go into modelling reflected fluxes and polarization fractions of Earth-like planet \citep{ treesstam2019,trees2022}.

The NASA Exoplanet Exploration Program’s Mission Star List for the Habitable Worlds Observatory includes an estimated contrast ratio for an Earth twin in the habitable zone for every star. 
%
The contrast is calculated as 
\begin{equation}
C = F_p/F_* = p \phi (\alpha) (R_p/a)^2,
\end{equation}
where p is the geometric albedo, $\phi (\alpha)$ is the integral phase function at phase angle $(\alpha)$, $R_p$ is the planet’s radius, and $a$ is the separation of the planet from its star. 
%
For their calculations they made the simplified assumptions of circular orbits, a geometric albedo of $p=0.2$, and that the scattering phase function, $\phi (\alpha)$, can be described by a simple Lambertian reflectance phase function, e.g. isotropic scattering.
%
These calculations give a good order-of-magnitude calculation for the contrast of Earth-like planets around nearby stars.
%
Therefore, in this work we will assume that the calculated contrasts given in the list correspond to the reflected flux at quadrature and normalize the scattering phase functions to this point.
%
We note that we do not aim to accurately calculate the absolute scaling of the contrast as this depends on many factors and would have required some of the work for the star list to be redone.

We use the scattering phase functions from \cite{treesstam2019} for an Earth-like planet with an ocean surface with patchy clouds and a wind-speed of $\SI{7}{\meter\per\second}$ at $\SI{670}{\nano\meter}$.
%
The $\SI{670}{\nano\meter}$ is close to the center of the Rc band for Vega ($\SI{642}{\nano\meter}$). 
%
We normalize the reflected light curve at the value of 90 degrees (quadrature) and multiply it with the contrast.
%
In addition, we map the orbital phase for circular orbits to the scattering phase angle and the on-sky separation given the inclinations of $60$ and $90$ degrees. \todo{Only the 60 degree plot is here at the moment}
%
Together they give the reflected light contrast as function of separation for a full orbit.
%
We multiply this flux with the degree of polarisation and retrieve the polarised contrast.
%
The resulting contrast curves are shown in \cref{fig:contrasts} for three of the closest target stars, the G dwarf $\alpha$ Cen A, the K dwarf $\epsilon$ Eri and the M dwarf Lalande 21185.

>>>>

@@

\subsection{Simulating contrast curves}
\textbf{Kim and David}
%
A second parameter that is crucial for detecting features like rainbows and ocean glint on Earth-like planets is the contrast, i.e. the relative flux of the reflected light from the planet divided by the stellar flux. 
%
While this is not the main focus of this paper, it is important to consider that the contrast and polarised contrast are not constant as function of scattering phase angles.
%
This is due to a combination of the changing illumination fraction of the exoplanets Earth facing surface and scattering effects.
%
Here, we aim to present some simplified examples of the impact of the scattering angle dependent contrast on the detectability of rainbow and glint features. 
%
We note that these calculations are by no means complete or exact, given the many variables that go into modelling reflected fluxes and polarization fractions of an Earth-like planet \citep{ treesstam2019,trees2022}.

The NASA Exoplanet Exploration Program’s Mission Star List for the Habitable Worlds Observatory includes an estimated contrast ratio for an Earth twin in the habitable zone for every star. 
%
The contrast is calculated as 
\begin{equation}
C = F_p/F_* = p \phi (\alpha) (R_p/a)^2,
\end{equation}
where p is the geometric albedo, $\phi (\alpha)$ is the integral phase function at phase angle $(\alpha)$, $R_p$ is the planet’s radius, and $a$ is the separation of the planet from its star. 
%
For their calculations they made the simplified assumptions of circular orbits, a geometric albedo of $p=0.2$, and that the scattering phase function, $\phi (\alpha)$, can be described by a simple Lambertian reflectance phase function, e.g. isotropic scattering.
%
These calculations give a good order-of-magnitude calculation for the contrast of Earth-like planets around nearby stars.
%
Therefore, in this work, we will assume that the calculated contrasts given in the list correspond to the reflected flux at quadrature and normalize the scattering phase functions to this point.
%
We note that we do not aim to accurately calculate the absolute scaling of the contrast, as this depends on many factors and would have required some of the work for the star list to be redone.

We use the scattering phase functions from \cite{treesstam2019} for an Earth-like planet with an ocean surface with patchy clouds and a wind-speed of $\SI{7}{\meter\per\second}$ at $\SI{670}{\nano\meter}$.
%
The $\SI{670}{\nano\meter}$ is close to the center of the Rc band for Vega ($\SI{642}{\nano\meter}$). 
%
We normalize the reflected light curve at the value of 90 degrees (quadrature) and multiply it with the contrast.
%
In addition, we map the orbital phase for circular orbits to the scattering phase angle and the on-sky separation given the inclinations of $60$ and $90$ degrees. \todo{Only the 60 degree plot is here at the moment}
%
Together they give the reflected light contrast as function of separation for a full orbit.
%
We multiply this flux with the degree of polarisation and retrieve the polarised contrast.
%
The resulting contrast curves are shown in \cref{fig:contrasts} for three of the closest target stars, the G~dwarf $\alpha$~Cen~A, the K~dwarf $\epsilon$~Eri and the M~dwarf Lalande~21185.
